Exploring dimensions of biodiversity in Japanese ferns

### 2020-07-02 ms/SI.Rmd
Add PD and RPE

ggsave(result_file("k-plot", "pdf"), width = 7, height = 3, units = "in")
```

@@

ggsave(result_file("k-plot-ecos", "pdf"), width = 4, height = 3, units = "in")
```

@@

```{r endemism-restricted, fig.width = 7, fig.height = 6}
# PE (endemic)
a <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = pe_obs)) +
  geom_sf(color = "transparent") +
  scale_fill_viridis_c() + 
  labs(fill = "PE") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme +
  theme(
    axis.text.x = element_blank()
  )

# SES PE (endemic)
b <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = pe_obs_z)) +
  geom_sf(color = "transparent") +
  labs(fill = "SES PE") +
  scale_fill_scico(
    palette = "romaO",
    na.value="grey80",
    direction = -1,
    limits = c(
      -get_limit(ses_div_ferns_endemic_spatial, pe_obs_z, "abs"),
      get_limit(ses_div_ferns_endemic_spatial, pe_obs_z, "abs")
    )) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme +
  theme(
    axis.text.y = element_blank()
  )

# CANAPE (endemic)
c <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = endem_type)) +
  geom_sf(color = "transparent") +
  labs(fill = "Endemism\ntype") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme

a + b + c + 
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A')

ggsave(result_file("endemism-restricted", "pdf"), width = 7, height = 6, units = "in")
```

**`r s_figure("endemism-restricted")`**. `r s_figure_cap("endemism-restricted")` **A**, phylogenetic endemism (PE). **B**, Standard effect size (SES) of PE. **C**, Categorical analysis of neo- and paleo-endemism (CANAPE).

@@

# Define helper function for making a scatterplot
scatterplot <- function(data, x_var, y_var, x_lab, y_lab, label, annotate) {
  p <- ggplot(data, aes(x = .data[[x_var]], y = .data[[y_var]])) +
    geom_point(alpha = 0.5) +
    labs(
      x = x_lab,
      y = y_lab
    )

  # Add annotation and trendline if annotate = TRUE
  if(annotate) p <- p +
      annotate("label", x = Inf, y = Inf, label = label, hjust = 1, vjust = 1, label.size = 0, fill = "transparent") +
      geom_smooth(method = "lm", se = FALSE, color = "blue", size = 0.5)

@@

```{r raw-biplot, fig.width = 7, fig.height = 6}
## Biplots of raw metrics ##

# Make dataframe of arguments to feed into plotting function
# - includes all non-duplicated combinations of raw diversity metrics
raw_comparisons <-
  cross_df(
    list(
      x_var = c("mpd_obs", "pe_obs", "mpd_morph_obs", "richness"),
      y_var = c("mpd_obs", "pe_obs", "mpd_morph_obs")
    ),
    .filter = `==` # filter out exact duplicates
  ) %>%
  # Remove reversed duplicates
  # https://stackoverflow.com/questions/22756392/deleting-reversed-duplicates-with-r
  group_by(grp = paste(pmax(x_var, y_var), pmin(x_var, y_var), sep = "_")) %>%
  # Use factors to keep the preferred combination of x and y
  mutate(x_var = as.factor(x_var) %>% fct_relevel("pe_obs", "mpd_morph_obs", "richness")) %>%
  mutate(y_var = as.factor(y_var) %>% fct_relevel("mpd_obs", "pe_obs", "mpd_morph_obs")) %>%
  arrange(x_var, y_var) %>%
  slice(1) %>%
  ungroup() %>%
  select(-grp) %>%
  # Convert x_var and y_var back to character for plotting function
  mutate(across(c(x_var, y_var), as.character)) %>%
  # Add labels
  mutate(
    x_lab = case_when(
      x_var == "mpd_morph_obs" ~ "MMD",
      x_var == "mpd_obs" ~ "MPD",
      x_var == "pe_obs" ~ "PE",
      x_var == "richness" ~ "Richness"
    ),
    y_lab = case_when(
      y_var == "mpd_morph_obs" ~ "MMD",
      y_var == "mpd_obs" ~ "MPD",
      y_var == "pe_obs" ~ "PE",
      y_var == "richness" ~ "Richness"
    )
  ) %>%
  # Run modified t-test
  mutate(
    test_res = map2(
      .x = x_var, 
      .y = y_var, 
      ~SpatialPack::modified.ttest(ses_div_ferns_spatial[[.x]], ses_div_ferns_spatial[[.y]], centroids)),
    # Extract relevant stats from results
    cor = map_dbl(test_res, pluck("corr")),
    pval = map_dbl(test_res, pluck("p.value"))
    ) %>%
  # Generate label
  mutate(
    cor_lab = glue("r = {round(cor, 3)}"),
    pval_lab = glue("p = {round(pval, 3)}"),
    label = glue::glue("{cor_lab}\n{ifelse(pval < 0.001, 'p < 0.001', pval_lab)}")
  ) %>%
  # Specify which plots to annotate
  mutate(
    annotate = case_when(
      x_var == "mpd_morph_obs" & y_var == "mpd_obs" & pval < 0.05 ~ TRUE,
      x_var == "mpd_morph_obs" & y_var == "pe_obs" & pval < 0.05 ~ TRUE,
      x_var == "mpd_obs" & y_var == "pe_obs" & pval < 0.05 ~ TRUE,
      x_var == "richness" & y_var == "pe_obs" & pval < 0.05  ~ TRUE,
      TRUE ~ FALSE
    )
  ) %>%
  select(x_var, y_var, x_lab, y_lab, label, annotate)

# Make plot
pmap(raw_comparisons, scatterplot, data = ses_div_ferns_spatial) %>%
wrap_plots() +
  plot_annotation(tag_levels = 'A')

ggsave(result_file("raw-biplot", "pdf"), width = 7, height = 6, units = "in")
```

**`r s_figure("raw-biplot")`**. `r s_figure_cap("raw-biplot")` MPD, mean phylogenetic distance; MMD, mean morphological distance; PE, phylogenetic endemism. Correlations between variables analyzed with Dutilleul's modified *t*-test that accounts for spatial structure in the data; *p* value and correlation (r) only shown for *p* < 0.05; correlations between MPD and MMD with richness not expected to be linear, so were not tested.

```{r ses-biplot, fig.width = 7, fig.height = 6}
## Biplots of SES metrics ##

# Make dataframe of arguments to feed into plotting function
# - includes all non-duplicated combinations of raw diversity metrics
ses_comparisons <-
  cross_df(
    list(
      x_var = c("mpd_obs_z", "pe_obs_z", "mpd_morph_obs_z", "richness"),
      y_var = c("mpd_obs_z", "pe_obs_z", "mpd_morph_obs_z")
    ),
    .filter = `==` # filter out exact duplicates
  ) %>%
  # Remove reversed duplicates
  # https://stackoverflow.com/questions/22756392/deleting-reversed-duplicates-with-r
  group_by(grp = paste(pmax(x_var, y_var), pmin(x_var, y_var), sep = "_")) %>%
  # Use factors to keep the preferred combination of x and y
  mutate(x_var = as.factor(x_var) %>% fct_relevel("pe_obs_z", "mpd_morph_obs_z", "richness")) %>%
  mutate(y_var = as.factor(y_var) %>% fct_relevel("mpd_obs_z", "pe_obs_z", "mpd_morph_obs_z")) %>%
  arrange(x_var, y_var) %>%
  slice(1) %>%
  ungroup() %>%
  select(-grp) %>%
  # Convert x_var and y_var back to character for plotting function
  mutate(across(c(x_var, y_var), as.character)) %>%
  # Add labels
  mutate(
    x_lab = case_when(
      x_var == "mpd_morph_obs_z" ~ "SES MMD",
      x_var == "mpd_obs_z" ~ "SES MPD",
      x_var == "pe_obs_z" ~ "SES PE",
      x_var == "richness" ~ "Richness"
    ),
    y_lab = case_when(
      y_var == "mpd_morph_obs_z" ~ "SES MMD",
      y_var == "mpd_obs_z" ~ "SES MPD",
      y_var == "pe_obs_z" ~ "SES PE",
      y_var == "richness" ~ "Richness"
    )
  ) %>%
  # Run modified t-test
  mutate(
    test_res = map2(
      .x = x_var, 
      .y = y_var, 
      ~SpatialPack::modified.ttest(ses_div_ferns_spatial[[.x]], ses_div_ferns_spatial[[.y]], centroids)),
    # Extract relevant stats from results
    cor = map_dbl(test_res, pluck("corr")),
    pval = map_dbl(test_res, pluck("p.value"))
    ) %>%
  # Generate label
  mutate(
    cor_lab = glue("r = {round(cor, 3)}"),
    pval_lab = glue("p = {round(pval, 3)}"),
    label = glue::glue("{cor_lab}\n{ifelse(pval < 0.001, 'p < 0.001', pval_lab)}")
  ) %>%
  # Specify which plots to annotate
  mutate(
    annotate = case_when(
      pval < 0.05 ~ TRUE,
      TRUE ~ FALSE
    )
  ) %>%
  select(x_var, y_var, x_lab, y_lab, label, annotate)

# Make plot
pmap(ses_comparisons, scatterplot, data = ses_div_ferns_spatial) %>%
  wrap_plots() +
  plot_annotation(tag_levels = 'A')

ggsave(result_file("ses-biplot", "pdf"), width = 7, height = 6, units = "in")
```

**`r s_figure("ses-biplot")`**. `r s_figure_cap("ses-biplot")` MPD, mean phylogenetic distance; MMD, mean morphological distance; PE, phylogenetic endemism. Correlations between variables analyzed with Dutilleul's modified *t*-test that accounts for spatial structure in the data; all combinations tested, but *p* value and correlation (r) only shown for *p* < 0.05.

@@

ggsave(result_file("traits-nmds", "pdf"), width = 8, height = 8, units = "in")
```

>>>>

ggsave(result_file(s_figure("k-plot"), "pdf"), width = 7, height = 3, units = "in")
```

@@

ggsave(result_file(s_figure("k-plot-ecos"), "pdf"), width = 4, height = 3, units = "in")
```

@@

```{r endemism-restricted, fig.width = 7, fig.height = 8}
# PE (endemic)
a <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = pe_obs)) +
  geom_sf(color = "transparent") +
  scale_fill_viridis_c() + 
  labs(fill = "PE") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme +
  theme(
    axis.text.x = element_blank()
  )

# SES PE (endemic)
b <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = pe_obs_z)) +
  geom_sf(color = "transparent") +
  labs(fill = "SES PE") +
  scale_fill_scico(
    palette = "romaO",
    na.value="grey80",
    direction = -1,
    limits = c(
      -get_limit(ses_div_ferns_endemic_spatial, pe_obs_z, "abs"),
      get_limit(ses_div_ferns_endemic_spatial, pe_obs_z, "abs")
    )) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme +
  theme(
    axis.text.y = element_blank(),
    axis.text.x = element_blank()
  )

# RPE
c <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = rpe_obs)) +
  geom_sf(color = "transparent") +
  scale_fill_viridis_c() + 
  labs(fill = "RPE") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme +
  theme(
    axis.text.x = element_blank()
  )

# SES RPE
d <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = rpe_obs_z)) +
  geom_sf(color = "transparent") +
  labs(fill = "SES RPE") +
  scale_fill_scico(
    palette = "romaO",
    na.value="grey80",
    direction = -1,
    limits = c(
      -get_limit(ses_div_ferns_endemic_spatial, rpe_obs_z, "abs"),
      get_limit(ses_div_ferns_endemic_spatial, rpe_obs_z, "abs")
    )) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme +
  theme(axis.text.y = element_blank())

# CANAPE (endemic)
e <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = endem_type)) +
  geom_sf(color = "transparent") +
  labs(fill = "Endemism\ntype") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme

a + b + c + d + e +
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A')

ggsave(result_file(s_figure("endemism-restricted"), "pdf"), width = 7, height = 8, units = "in")
```

**`r s_figure("endemism-restricted")`**. `r s_figure_cap("endemism-restricted")` **A**, phylogenetic endemism (PE). **B**, Standard effect size (SES) of PE. **C**, relative phylogenetic endemism (RPE). **D**, SES of RPE. **E**, Categorical analysis of neo- and paleo-endemism (CANAPE).

@@

# Define helper function for making a scatterplot
scatterplot <- function(data, x_var, y_var, x_lab, y_lab, label, annotate, loc) {
  p <- ggplot(data, aes(x = .data[[x_var]], y = .data[[y_var]])) +
    geom_point(alpha = 0.5) +
    labs(
      x = x_lab,
      y = y_lab
    )

  # Add annotation and trendline if annotate = TRUE
  if(annotate & loc == "UL") p <- p +
      annotate("label", x = -Inf, y = Inf, label = label, hjust = 0, vjust = 1, label.size = 0, size = 3, fill = "transparent") +
      geom_smooth(method = "lm", se = FALSE, color = "blue", size = 0.5)

  if(annotate & loc == "UR") p <- p +
      annotate("label", x = Inf, y = Inf, label = label, hjust = 1, vjust = 1, label.size = 0, size = 3, fill = "transparent") +
      geom_smooth(method = "lm", se = FALSE, color = "blue", size = 0.5)

  if(annotate & loc == "LL") p <- p +
      annotate("label", x = -Inf, y = -Inf, label = label, hjust = 0, vjust = 0, label.size = 0, size = 3, fill = "transparent") +
      geom_smooth(method = "lm", se = FALSE, color = "blue", size = 0.5)

  if(annotate & loc == "LR") p <- p +
      annotate("label", x = Inf, y = -Inf, label = label, hjust = 1, vjust = 0, label.size = 0, size = 3, fill = "transparent") +
      geom_smooth(method = "lm", se = FALSE, color = "blue", size = 0.5)

@@

```{r raw-biplot, fig.width = 8, fig.height = 10}
## Biplots of raw metrics ##

# Set up x and y variables as factors so they can be ordered appropriately in the plot
x_vars <-
  c("mpd_obs", "pd_obs", "pe_obs", "rpe_obs", "mpd_morph_obs", "richness") %>%
  factor() %>%
  fct_relevel(as.character(.))

y_vars <- c("mpd_obs", "pd_obs", "pe_obs", "rpe_obs", "mpd_morph_obs") %>%
  factor() %>%
  fct_relevel(as.character(.))

# Make dataframe of arguments to feed into plotting function
# - includes all non-duplicated combinations of raw diversity metrics
raw_comparisons <-
  cross_df(
    list(
      x_var = x_vars,
      y_var = y_vars
    ),
    .filter = `==` # filter out exact duplicates
  ) %>%
  # Remove reversed duplicates
  # https://stackoverflow.com/questions/22756392/deleting-reversed-duplicates-with-r
  group_by(grp = paste(pmax(as.character(x_var), as.character(y_var)), pmin(as.character(x_var), as.character(y_var)), sep = "_")) %>%
  arrange(x_var, y_var) %>%
  slice(1) %>%
  ungroup() %>%
  select(-grp) %>%
  arrange(x_var, y_var) %>%
  # Convert x_var and y_var back to character for plotting function
  mutate(across(c(x_var, y_var), as.character)) %>%
  # Add labels
  mutate(
    x_lab = case_when(
      x_var == "mpd_morph_obs" ~ "MMD",
      x_var == "mpd_obs" ~ "MPD",
      x_var == "pd_obs" ~ "PD",
      x_var == "rpe_obs" ~ "RPE",
      x_var == "pe_obs" ~ "PE",
      x_var == "richness" ~ "Richness"
    ),
    y_lab = case_when(
      y_var == "mpd_morph_obs" ~ "MMD",
      y_var == "mpd_obs" ~ "MPD",
      y_var == "pd_obs" ~ "PD",
      y_var == "rpe_obs" ~ "RPE",
      y_var == "pe_obs" ~ "PE",
      y_var == "richness" ~ "Richness"
    )
  ) %>%
  # Run modified t-test
  mutate(
    test_res = map2(
      .x = x_var, 
      .y = y_var, 
      ~SpatialPack::modified.ttest(ses_div_ferns_spatial[[.x]], ses_div_ferns_spatial[[.y]], centroids)),
    # Extract relevant stats from results
    cor = map_dbl(test_res, pluck("corr")),
    pval = map_dbl(test_res, pluck("p.value"))
    ) %>%
  # Generate label
  mutate(
    cor_lab = glue("r = {round(cor, 3)}"),
    pval_lab = glue("p = {round(pval, 3)}"),
    label = glue::glue("{cor_lab}\n{ifelse(pval < 0.001, 'p < 0.001', pval_lab)}")
  ) %>%
  # Specify which plots to annotate
  mutate(
    annotate = case_when(
      pval < 0.05 ~ TRUE,
      TRUE ~ FALSE
    ),
    # annotate in upper-right corner for mpd_morph_obs + rpe_obs,
    # upper-left otherwise
    loc = case_when(
      y_var == "mpd_morph_obs" & x_var == "rpe_obs" ~ "UR",
      TRUE ~ "UL"
    )
  ) %>%
  select(x_var, y_var, x_lab, y_lab, label, annotate, loc)

# Make plot
pmap(raw_comparisons, scatterplot, data = ses_div_ferns_spatial) %>%
wrap_plots() +
  plot_layout(ncol = 3) +
  plot_annotation(tag_levels = 'A')

ggsave(result_file(s_figure("raw-biplot"), "pdf"), width = 8, height = 10, units = "in")
```

**`r s_figure("raw-biplot")`**. `r s_figure_cap("raw-biplot")` MPD, mean phylogenetic distance; MMD, mean morphological distance; PD, phylogenetic diversity; PE, phylogenetic endemism; RPE, relative phylogenetic endemism. Correlations between variables analyzed with Dutilleul's modified *t*-test that accounts for spatial structure in the data; *p* value and correlation (r) only shown for *p* < 0.05.

```{r ses-biplot, fig.width = 8, fig.height = 10}
## Biplots of SES metrics ##

# Set up x and y variables as factors so they can be ordered appropriately in the plot
x_vars <-
  c("mpd_obs_z", "pd_obs_z", "pe_obs_z", "rpe_obs_z", "mpd_morph_obs_z", "richness") %>%
  factor() %>%
  fct_relevel(as.character(.))

y_vars <- c("mpd_obs_z", "pd_obs_z", "pe_obs_z", "rpe_obs_z", "mpd_morph_obs_z") %>%
  factor() %>%
  fct_relevel(as.character(.))

# Make dataframe of arguments to feed into plotting function
# - includes all non-duplicated combinations of raw diversity metrics
ses_comparisons <-
  cross_df(
    list(
      x_var = x_vars,
      y_var = y_vars
    ),
    .filter = `==` # filter out exact duplicates
  ) %>%
  # Remove reversed duplicates
  # https://stackoverflow.com/questions/22756392/deleting-reversed-duplicates-with-r
  group_by(grp = paste(pmax(as.character(x_var), as.character(y_var)), pmin(as.character(x_var), as.character(y_var)), sep = "_")) %>%
  arrange(x_var, y_var) %>%
  slice(1) %>%
  ungroup() %>%
  select(-grp) %>%
  arrange(x_var, y_var) %>%
  # Convert x_var and y_var back to character for plotting function
  mutate(across(c(x_var, y_var), as.character)) %>%
  # Add labels
  mutate(
    x_lab = case_when(
      x_var == "mpd_morph_obs_z" ~ "SES MMD",
      x_var == "mpd_obs_z" ~ "SES MPD",
      x_var == "pd_obs_z" ~ "SES PD",
      x_var == "pe_obs_z" ~ "SES PE",
      x_var == "rpe_obs_z" ~ "SES RPE",
      x_var == "richness" ~ "Richness"
    ),
    y_lab = case_when(
      y_var == "mpd_morph_obs_z" ~ "SES MMD",
      y_var == "mpd_obs_z" ~ "SES MPD",
      y_var == "pd_obs_z" ~ "SES PD",
      y_var == "pe_obs_z" ~ "SES PE",
      y_var == "rpe_obs_z" ~ "SES RPE",
      y_var == "richness" ~ "Richness"
    )
  ) %>%
  # Run modified t-test
  mutate(
    test_res = map2(
      .x = x_var, 
      .y = y_var, 
      ~SpatialPack::modified.ttest(ses_div_ferns_spatial[[.x]], ses_div_ferns_spatial[[.y]], centroids)),
    # Extract relevant stats from results
    cor = map_dbl(test_res, pluck("corr")),
    pval = map_dbl(test_res, pluck("p.value"))
    ) %>%
  # Generate label
  mutate(
    cor_lab = glue("r = {round(cor, 3)}"),
    pval_lab = glue("p = {round(pval, 3)}"),
    label = glue::glue("{cor_lab}\n{ifelse(pval < 0.001, 'p < 0.001', pval_lab)}")
  ) %>%
  # Specify which plots to annotate
  mutate(
    annotate = case_when(
      pval < 0.05 ~ TRUE,
      TRUE ~ FALSE
    ),
    # annotate in upper-right corner for mpd_morph_obs + rpe_obs,
    # upper-left otherwise
    loc = case_when(
      y_var == "mpd_morph_obs_z" & x_var == "pd_obs_z" ~ "LR",
      y_var == "rpe_obs_z" & x_var == "pe_obs_z" ~ "LR",
      y_var == "mpd_morph_obs_z" & x_var == "pe_obs_z" ~ "LR",
      y_var == "pd_obs_z" & x_var == "richness" ~ "UR",
      TRUE ~ "UL"
    )
  ) %>%
  select(x_var, y_var, x_lab, y_lab, label, annotate, loc)

# Make plot
pmap(ses_comparisons, scatterplot, data = ses_div_ferns_spatial) %>%
wrap_plots() +
  plot_layout(ncol = 3) +
  plot_annotation(tag_levels = 'A')

ggsave(result_file(s_figure("ses-biplot"), "pdf"), width = 8, height = 10, units = "in")
```

**`r s_figure("ses-biplot")`**. `r s_figure_cap("ses-biplot")` MPD, mean phylogenetic distance; MMD, mean morphological distance; PD, phylogenetic diversity; PE, phylogenetic endemism; RPE, relative phylogenetic endemism. Correlations between variables analyzed with Dutilleul's modified *t*-test that accounts for spatial structure in the data; *p* value and correlation (r) only shown for *p* < 0.05.

@@

ggsave(result_file(s_figure("traits-nmds"), "pdf"), width = 8, height = 8, units = "in")
```

### 2020-07-02 ms/manuscript.Rmd
Add PD and RPE

Joel H. Nitta^1^ and Atsushi Ebihara^2^

@@

^2^Department of Botany, National Museum of Nature and Science, 4-1-1 Amakubo, Tsukuba-shi, Ibaraki, 305-0005, Japan

@@

Raw values of MPD and MMD showed high variance and low species richness, then merged towards a single value at high species richness, as expected (`r s_figure("raw-biplot")`D, E). Raw PE was linearly correlated with species richness (`r s_figure("raw-biplot")`F). Both raw and SES values of MPD and MMD were highly correlated (`r s_figure("raw-biplot")`A, `r s_figure("ses-biplot")`A, B). SES values of MPD, MMD, and PE were all positively correlated (`r s_figure("ses-biplot")`A, B, C). None of the SES values were correlated with richness (`r s_figure("ses-biplot")`D, E, F).

@@

# Raw morph diversity
e <- ggplot(ses_div_ferns_spatial, aes(fill = mpd_morph_obs)) +
  geom_sf(color = "transparent") +
  labs(fill = "MMD") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  scale_fill_viridis_c() +
  map_theme

# SES of morph diversity
f <- ggplot(ses_div_ferns_spatial, aes(fill = mpd_morph_obs_z)) +
  geom_sf(color = "transparent") +
  labs(fill = "SES MMD") +
  scale_fill_scico(
    palette = "romaO",
    na.value="grey80",
    direction = -1,
    limits = c(
      -get_limit(ses_div_ferns_spatial, mpd_morph_obs_z, "abs"),
      get_limit(ses_div_ferns_spatial, mpd_morph_obs_z, "abs")
    )) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme +
  theme(
    axis.text.y = element_blank()
  )

a + b + c + d + e + f +
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A')

ggsave(here::here("results/Fig_1.pdf"), width = 7, height = 9, units = "in")
```

**`r figure("diversity")`**. `r figure_cap("diversity")` **A**, richness. **B**, sampling redundancy. **C**, mean phylogenetic distance (MPD). **D**, Standard effect size (SES) of MPD. **E**, mean morphological distance (MMD) **F**, SES of MMD.

@@

```{r endemism, fig.width = 7, fig.height = 6}
# PE
a <- ggplot(ses_div_ferns_spatial, aes(fill = pe_obs)) +
  geom_sf(color = "transparent") +
  scale_fill_viridis_c() + 
  labs(fill = "PE") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme +
  theme(
    axis.text.x = element_blank()
  )

@@

# CANAPE
c <- ggplot(ses_div_ferns_spatial, aes(fill = endem_type)) +
  geom_sf(color = "transparent") +
  labs(fill = "Endemism\ntype") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme

a + b + c +
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A')

ggsave(here::here("results/Fig_2.pdf"), width = 7, height = 6, units = "in")
```

**`r figure("endemism")`**. `r figure_cap("endemism")` **A**, phylogenetic endemism (PE). **B**, Standard effect size (SES) of PE. **C**, Categorical analysis of neo- and paleo-endemism (CANAPE).

@@

ggsave(here::here("results/Fig_3.pdf"), width = 7, height = 3, units = "in")
```

>>>>

Joel H. Nitta^1^, Brent D. Mishler^2^, Wataru Iwasaki^1^, and Atsushi Ebihara^3^

@@

^2^University and Jepson Herbaria, University of California, Berkeley, CA 94720‐2465, USA

^3^Department of Botany, National Museum of Nature and Science, 4-1-1 Amakubo, Tsukuba-shi, Ibaraki, 305-0005, Japan

@@

Raw values of MPD and MMD showed high variance and low species richness, then merged towards a single value at high species richness, as expected (`r s_figure("raw-biplot")`K, O). Raw PE was linearly correlated with species richness (`r s_figure("raw-biplot")`M). Both raw and SES values of MPD and MMD were highly correlated (`r s_figure("raw-biplot")`D, `r s_figure("ses-biplot")`D). SES values of MPD, MMD, and PE were all positively correlated (`r s_figure("ses-biplot")`). None of the SES values were correlated with richness (`r s_figure("ses-biplot")`).

@@

# Raw morph diversity
e <- ggplot(ses_div_ferns_spatial, aes(fill = mpd_morph_obs)) +
  geom_sf(color = "transparent") +
  labs(fill = "MMD") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  scale_fill_viridis_c() +
  map_theme +
  theme(
    axis.text.x = element_blank()
  )

# SES of morph diversity
f <- ggplot(ses_div_ferns_spatial, aes(fill = mpd_morph_obs_z)) +
  geom_sf(color = "transparent") +
  labs(fill = "SES MMD") +
  scale_fill_scico(
    palette = "romaO",
    na.value="grey80",
    direction = -1,
    limits = c(
      -get_limit(ses_div_ferns_spatial, mpd_morph_obs_z, "abs"),
      get_limit(ses_div_ferns_spatial, mpd_morph_obs_z, "abs")
    )) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank()
  )

g <- ggplot(ses_div_ferns_spatial, aes(fill = pd_obs)) +
  geom_sf(color = "transparent") +
  labs(fill = "MMD") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  scale_fill_viridis_c() +
  map_theme

h <- ggplot(ses_div_ferns_spatial, aes(fill = pd_obs_z)) +
  geom_sf(color = "transparent") +
  labs(fill = "SES PD") +
  scale_fill_scico(
    palette = "romaO",
    na.value="grey80",
    direction = -1,
    limits = c(
      -get_limit(ses_div_ferns_spatial, pd_obs_z, "abs"),
      get_limit(ses_div_ferns_spatial, pd_obs_z, "abs")
    )) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme +
  theme(
    axis.text.y = element_blank()
  )

a + b + c + d + e + f + g + h +
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A')

ggsave(result_file(figure("diversity"), "pdf"), width = 7, height = 9, units = "in")
```

**`r figure("diversity")`**. `r figure_cap("diversity")` **A**, richness. **B**, sampling redundancy. **C**, mean phylogenetic distance (MPD). **D**, Standard effect size (SES) of MPD. **E**, mean morphological distance (MMD) **F**, SES of MMD. **G**, phylogenetic diversity (PD) **H**, SES of PD.

@@

```{r endemism, fig.width = 7, fig.height = 8}
# PE
a <- ggplot(ses_div_ferns_spatial, aes(fill = pe_obs)) +
  geom_sf(color = "transparent") +
  scale_fill_viridis_c() + 
  labs(fill = "PE") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme +
  theme(
    axis.text.x = element_blank()
  )

@@

# RPE
c <- ggplot(ses_div_ferns_spatial, aes(fill = rpe_obs)) +
  geom_sf(color = "transparent") +
  scale_fill_viridis_c() + 
  labs(fill = "RPE") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme +
  theme(
    axis.text.x = element_blank()
  )

# SES RPE
d <- ggplot(ses_div_ferns_spatial, aes(fill = rpe_obs_z)) +
  geom_sf(color = "transparent") +
  labs(fill = "SES RPE") +
  scale_fill_scico(
    palette = "romaO",
    na.value="grey80",
    direction = -1,
    limits = c(
      -get_limit(ses_div_ferns_spatial, rpe_obs_z, "abs"),
      get_limit(ses_div_ferns_spatial, rpe_obs_z, "abs")
    )) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme +
  theme(
    axis.text.y = element_blank()
  )

# CANAPE
e <- ggplot(ses_div_ferns_spatial, aes(fill = endem_type)) +
  geom_sf(color = "transparent") +
  labs(fill = "Endemism\ntype") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme

a + b + c + d + e +
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A')

ggsave(result_file(figure("endemism"), "pdf"), width = 7, height = 8, units = "in")
```

**`r figure("endemism")`**. `r figure_cap("endemism")` **A**, phylogenetic endemism (PE). **B**, Standard effect size (SES) of PE. **C**, relative phylogenetic endemism (RPE). **D**, SES of RPE. **E**, Categorical analysis of neo- and paleo-endemism (CANAPE).

@@

ggsave(result_file(figure("bioregions"), "pdf"), width = 7, height = 3, units = "in")
```
